\textit{Red Oak}\\ Proactive Controller Generation for Algorithmic Policy

### 2015-05-12 draft.tex
fixed ipsrc ipdst to Src and Dst

\lstset{morekeywords={in_range, in_rel, add, remove, Drop, Deliver, ForwardTo, Multicast, IpSrc, IpDst}}

@@

\begin{figure}[ht]
\begin{lstlisting}
f(pkt):
  if in_range(pkt, src, 0, 10) then
     if in_range pkt dst  [(0,10)] then
        Deliver
     else 
        add pkt (IpDst, IpSrc) ``conn'';
        Deliver
  else if in_rel pkt (IpSrc,IpDst) ``conn'' then
         Deliver
       else
         Drop
       end
  end                
  \end{lstlisting}

@@

This decision tree would, in turn, lead to the following set of prioritized symbolic rules. We call these symbolic rules because they represent constraints on the actual concrete rules. The concrete rules our system will install in the network depend on the current contents of the trusted connection relation. In particular, for each IP source that is trusted to communicate with an internal destination in the "conn" relation, we will get a concrete rule permitting this communication.
  \begin{lstlisting}[mathescape]
src $\in$ [0, 100] $\mapsto$ Controller
src $\in$ [101, 1000], (src,dst) $\in$ conn  $\mapsto$ Deliver
src $\in$ [101, 1000], (src,dst) $\notin$ conn $\mapsto$ Drop
  \end{lstlisting}

After initialization (starting with no trusted connections), the concrete rules would be generated from the symbolic rules by having all relations be empty:
\begin{lstlisting}[mathescape]
src $\in$ [0, 100] $\mapsto$ Controller  
src $\in$ [101, 1000] $\mapsto$ Drop  
\end{lstlisting}

@@

  For example, a packing coming from \lstinline|src| 150 and \lstinline|dst| 10 would be matched by the second rule, and dropped on the switch.
Subsequently, on a packet with \lstinline|src| 10 and \lstinline|dst| 150, the first rules would be matched, evaluating the decision tree would add tuple $(10,150)$ to relation \lstinline|conn|, the packet would be forwarded according to the forwarding decision \lstinline|Deliver|, and the concrete rules would be updated as follows:

\begin{lstlisting}[mathescape]
src $\in$ [0, 100] $\mapsto$ Controller  
src = 150, dst = 10 $\mapsto$ Deliver  
src $\in$ [101, 1000] $\mapsto$ Drop  
\end{lstlisting}

@@

   \subsection*{Forwarding Decision}
   % can give constant decision (e.g. drop() or fwd(loc) to a certain location loc), or decision based on values of the packet (e.g. deliver() which forwards to dst)
	The programming model for \textit{Red Oak} is that each packet entering the network is mapped to a forwarding decision that describes where the packet should end up leaving the network. For simplicity, we assume that we are targeting a one big switch implementation of the network. The one big switch model has been heavily optimized (e.g., \cite{Obs}), and we view the problem of determining paths through the network as orthogonal to our work here.

@@

let filter_src_range f r pkt =
 if in_range pkt IpSrc r then
    f pkt
 else
    Drop   
\end{lstlisting}

@@

   For example, we include in Figure~\ref{fig:ex-hole} the source code for \lstinline|filter_src_range| a partial policy, which, given a policy \lstinline|f| and a range \lstinline|r|, will forward all packets whose source IP is within range \lstinline|r| and drop the packet otherwise. The ability to compose two or more policies in sequence like this arises many times in practice. In our earlier firewall example, instead of delivering packets that were not dropped, we could instead have deferred to another function to decide how to handle trusted traffic.

@@

\begin{figure}[ht]
  \begin{lstlisting}[mathescape]
type pkt
type field = IpSrc | IpDst 
type fields = field list
type range = (int*int) list

@@

For example, taking the following symbolic rules from our earlier stateful firewall example: 
  \begin{lstlisting}[mathescape]
src $\in$ [101, 1000], (src,dst) $\in$ conn $\mapsto$ Deliver
src $\in$ [101, 1000], (src,dst) $\notin$ conn $\mapsto$ Drop
  \end{lstlisting}
The first rules says: "external hosts that are trusted should be let through", and the second rules says: "traffic from untrusted external hosts should be dropped".

  Assume that the trusted connection relation ``conn'' contains the pairs 
  \lstinline| {(150,10), (200, 9), (8, 150)}|, meaning external src 150 can talk to internal dst 10 and so on. We concretize the above symbolic rules by expanding the first rule to:
\begin{lstlisting}[mathescape]
src $\in$ ([101, 1000] $\cap$ 150), dst = 10 $\mapsto$ Deliver
src $\in$ ([101, 1000] $\cap$ 200), dst = 9 $\mapsto$ Deliver
src $\in$ ([101, 1000] $\cap$ 8), dst = 150 $\mapsto$ Deliver
 \end{lstlisting}
To satisfy both symbolic constraints, we take the intersection of the constraints, which corresponds to intersection of the constraints on the ranges of each field. These rules reduce to:
\begin{lstlisting}[mathescape]
src = 150, dst = 10 $\mapsto$ Deliver
src = 200, dst = 9 $\mapsto$ Deliver
 \end{lstlisting}
Note the omission of the third concrete rule, resulting from \lstinline|ipsrc| having an empty range. We can now concretize the second symbolic rule without doing any work, since checking for relation membership is already done by all the higher priority rules. The final concrete forwarding rules are now:
\begin{lstlisting}[mathescape]
src = 150, dst = 10 $\mapsto$ Deliver
src = 200, dst = 9 $\mapsto$ Deliver
src $\in$ [101, 1000] $\mapsto$ Drop
\end{lstlisting}

@@

\subsection*{Optimizations}
   % 1) expand tree before add/remove nodes with a match on that relation ( based on add(p); add(p) == add(p) so can shortcircuit if p is already in the relation)
The rules generated by the presented technique ensures that the only packets being forwarded to the controller are those reaching an add or a remove node and thus potentially modify the state of the controller. However, these packets do not always modify the state of the controller. For example, in the firewall example (see Fig~\ref{fig:firewallcode}), after seeing a first packet with an \lstinline|ipsrc| of 10 and \lstinline|ipdst| of 150, any subsequent packet with the same \lstinline|ipsrc| and \lstinline|ipdst| could be delivered without being first forwarded to the controller, as adding the same tuple to a relation is idempotent. By using both local and global transformations on the decision tree, we can get closer to an optimal number of packet forwarded to the controller, where only packets having a real, meaningful effect are seen by the controller.

@@

From which the following symbolic rule would be generate:
\begin{lstlisting}[mathescape]
src $\in$ [0, 100], (dst, src) in conn $\mapsto$ Deliver
src $\in$ [0, 100] $\mapsto$ Controller
src $\in$ [101, 1000], (src,dst) $\in$ conn $\mapsto$ Deliver
src $\in$ [101, 1000], (src,dst) $\notin$ conn $\mapsto$ Drop
\end{lstlisting}

With empty relation, the concrete rules would be the same as with the unoptimized decision tree. However, after seeing, for example, packet with \lstinline|ipsrc| 10 and \lstinline|ipdst| 150, the rules would be
\begin{lstlisting}[mathescape]
src = 10, dst = 150 $\mapsto$ Deliver  
src $\in$ [0, 100] $\mapsto$ Controller  
src = 150, dst = 10 $\mapsto$ Deliver  
src $\in$ [101, 1000] $\mapsto$ Drop  
\end{lstlisting}

Other optimizations may further reduce the number of packet sent to the controller. For example, we could remove \lstinline|add| node which are followed by \lstinline|remove| node for the same relation and tuple on all path without the change being observed by an \lstinline|in_rel| node. Similarly, we could detect and warn the user if an added tuple may never be observed by an \lstinline|in_rel| node, for example by switching \lstinline|ipsrc| and \lstinline|ipdst| in the add tuple of the stateful firewall example (see Figure~\ref{fig:firewallcode}). We leave the implementation of these optimizations as future work.

>>>>

\lstset{morekeywords={in_range, in_rel, add, remove, Drop, Deliver, ForwardTo, Multicast, Src, Dst}}

@@

\begin{figure}[ht]
\begin{lstlisting}
f(pkt):
  if in_range(pkt, Src, 0, 10) then
     if in_range pkt Dst  [(0,10)] then
        Deliver
     else 
        add pkt (Dst, Src) ``conn'';
        Deliver
  else if in_rel pkt (Src,Dst) ``conn'' then
         Deliver
       else
         Drop
       end
  end                
  \end{lstlisting}

@@

This decision tree would, in turn, lead to the following set of prioritized symbolic rules. We call these symbolic rules because they represent constraints on the actual concrete rules. The concrete rules our system will install in the network depend on the current contents of the trusted connection relation. In particular, for each IP source that is trusted to communicate with an internal destination in the "conn" relation, we will get a concrete rule permitting this communication.
  \begin{lstlisting}[mathescape]
Src $\in$ [0, 100] $\mapsto$ Controller
Src $\in$ [101, 1000], (Src,Dst) $\in$ conn  $\mapsto$ Deliver
Src $\in$ [101, 1000], (Src,Dst) $\notin$ conn $\mapsto$ Drop
  \end{lstlisting}

After initialization (starting with no trusted connections), the concrete rules would be generated from the symbolic rules by having all relations be empty:
\begin{lstlisting}[mathescape]
Src $\in$ [0, 100] $\mapsto$ Controller  
Src $\in$ [101, 1000] $\mapsto$ Drop  
\end{lstlisting}

@@

  For example, a packing coming from \lstinline|Src| 150 and \lstinline|Dst| 10 would be matched by the second rule, and dropped on the switch.
Subsequently, on a packet with \lstinline|Src| 10 and \lstinline|Dst| 150, the first rules would be matched, evaluating the decision tree would add tuple $(10,150)$ to relation \lstinline|conn|, the packet would be forwarded according to the forwarding decision \lstinline|Deliver|, and the concrete rules would be updated as follows:

\begin{lstlisting}[mathescape]
Src $\in$ [0, 100] $\mapsto$ Controller  
Src = 150, Dst = 10 $\mapsto$ Deliver  
Src $\in$ [101, 1000] $\mapsto$ Drop  
\end{lstlisting}

@@

   \subsection*{Forwarding Decision}
   % can give constant decision (e.g. drop() or fwd(loc) to a certain location loc), or decision based on values of the packet (e.g. deliver() which forwards to Dst)
	The programming model for \textit{Red Oak} is that each packet entering the network is mapped to a forwarding decision that describes where the packet should end up leaving the network. For simplicity, we assume that we are targeting a one big switch implementation of the network. The one big switch model has been heavily optimized (e.g., \cite{Obs}), and we view the problem of determining paths through the network as orthogonal to our work here.

@@

let filter_Src_range f r pkt =
 if in_range pkt Src r then
    f pkt
 else
    Drop   
\end{lstlisting}

@@

   For example, we include in Figure~\ref{fig:ex-hole} the source code for \lstinline|filter_Src_range| a partial policy, which, given a policy \lstinline|f| and a range \lstinline|r|, will forward all packets whose source IP is within range \lstinline|r| and drop the packet otherwise. The ability to compose two or more policies in sequence like this arises many times in practice. In our earlier firewall example, instead of delivering packets that were not dropped, we could instead have deferred to another function to decide how to handle trusted traffic.

@@

\begin{figure}[ht]
  \begin{lstlisting}[mathescape]
type pkt
type field = Src | Dst 
type fields = field list
type range = (int*int) list

@@

For example, taking the following symbolic rules from our earlier stateful firewall example: 
  \begin{lstlisting}[mathescape]
Src $\in$ [101, 1000], (Src, Dst) $\in$ conn $\mapsto$ Deliver
Src $\in$ [101, 1000], (Src, Dst) $\notin$ conn $\mapsto$ Drop
  \end{lstlisting}
The first rules says: "external hosts that are trusted should be let through", and the second rules says: "traffic from untrusted external hosts should be dropped".

  Assume that the trusted connection relation ``conn'' contains the pairs 
  \lstinline| {(150,10), (200, 9), (8, 150)}|, meaning external Src 150 can talk to internal Dst 10 and so on. We concretize the above symbolic rules by expanding the first rule to:
\begin{lstlisting}[mathescape]
Src $\in$ ([101, 1000] $\cap$ 150), Dst = 10 $\mapsto$ Deliver
Src $\in$ ([101, 1000] $\cap$ 200), Dst = 9 $\mapsto$ Deliver
Src $\in$ ([101, 1000] $\cap$ 8), Dst = 150 $\mapsto$ Deliver
 \end{lstlisting}
To satisfy both symbolic constraints, we take the intersection of the constraints, which corresponds to intersection of the constraints on the ranges of each field. These rules reduce to:
\begin{lstlisting}[mathescape]
Src = 150, Dst = 10 $\mapsto$ Deliver
Src = 200, Dst = 9 $\mapsto$ Deliver
 \end{lstlisting}
Note the omission of the third concrete rule, resulting from \lstinline|Src| having an empty range. We can now concretize the second symbolic rule without doing any work, since checking for relation membership is already done by all the higher priority rules. The final concrete forwarding rules are now:
\begin{lstlisting}[mathescape]
Src = 150, Dst = 10 $\mapsto$ Deliver
Src = 200, Dst = 9 $\mapsto$ Deliver
Src $\in$ [101, 1000] $\mapsto$ Drop
\end{lstlisting}

@@

\subsection*{Optimizations}
   % 1) expand tree before add/remove nodes with a match on that relation ( based on add(p); add(p) == add(p) so can shortcircuit if p is already in the relation)
The rules generated by the presented technique ensures that the only packets being forwarded to the controller are those reaching an add or a remove node and thus potentially modify the state of the controller. However, these packets do not always modify the state of the controller. For example, in the firewall example (see Fig~\ref{fig:firewallcode}), after seeing a first packet with an \lstinline|Src| of 10 and \lstinline|Dst| of 150, any subsequent packet with the same \lstinline|Src| and \lstinline|Dst| could be delivered without being first forwarded to the controller, as adding the same tuple to a relation is idempotent. By using both local and global transformations on the decision tree, we can get closer to an optimal number of packet forwarded to the controller, where only packets having a real, meaningful effect are seen by the controller.

@@

From which the following symbolic rule would be generate:
\begin{lstlisting}[mathescape]
Src $\in$ [0, 100], (Dst, Src) in conn $\mapsto$ Deliver
Src $\in$ [0, 100] $\mapsto$ Controller
Src $\in$ [101, 1000], (Src,Dst) $\in$ conn $\mapsto$ Deliver
Src $\in$ [101, 1000], (Src,Dst) $\notin$ conn $\mapsto$ Drop
\end{lstlisting}

With empty relation, the concrete rules would be the same as with the unoptimized decision tree. However, after seeing, for example, packet with \lstinline|Src| 10 and \lstinline|Dst| 150, the rules would be
\begin{lstlisting}[mathescape]
Src = 10, Dst = 150 $\mapsto$ Deliver  
Src $\in$ [0, 100] $\mapsto$ Controller  
Src = 150, Dst = 10 $\mapsto$ Deliver  
Src $\in$ [101, 1000] $\mapsto$ Drop  
\end{lstlisting}

Other optimizations may further reduce the number of packet sent to the controller. For example, we could remove \lstinline|add| node which are followed by \lstinline|remove| node for the same relation and tuple on all path without the change being observed by an \lstinline|in_rel| node. Similarly, we could detect and warn the user if an added tuple may never be observed by an \lstinline|in_rel| node, for example by switching \lstinline|Src| and \lstinline|Dst| in the add tuple of the stateful firewall example (see Figure~\ref{fig:firewallcode}). We leave the implementation of these optimizations as future work.
